---
title: "a3task3_ea"
author: "Elmera Azadpour"
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

## Call in urchins data and clean column names
```{r, message = FALSE, warning=FALSE}
urchins <- read_csv(here("data","urchins.csv"))
urchins_clean <- urchins %>% clean_names()
```

## i. Exploratory data visualizaton
```{r, message = FALSE, warning=FALSE}
ggplot(data=urchins_clean, aes(x=size)) +
  geom_histogram(bins = 15) +
  facet_wrap(~common_name)

ggplot(data=urchins_clean, aes(x=common_name, y=size)) +
  geom_boxplot()
```

## ii. Descriptive statistics table
```{r, message = FALSE, warning=FALSE}
urchins_stats <- urchins_clean %>% 
  group_by(common_name) %>% 
  summarise(mean_size = mean(size),
            median_size = median((size)),
            sd_size = sd(size),
            n = n())

urchins_stats %>% 
  kable(col.names = c("Species",
                      "Mean size (cm)", 
                      "Median size (cm)",
                      "Standard deviation",
                      "Sample size (n)")) %>% 
  kable_styling(full_width = FALSE)
```

## iii. sentences briefly describing your overall takeaways about how the size (distribution, central tendency, spread, etc.) compare between the two species of urchins:
- Looking at the histogram, the red urchins data seems to be more normally distributed whereas the purple urchins data seems to be skewed a little more towards smaller sizes. This follows the trend because the mean for purple is 2.31 less than red. Purple has a smaller standard deviation suggesting that the values are closer to the mean, whereas red has a larger sd suggesting that the values are a bit more spread. It should also be noted that purple urchins sample size is about 4000 greater than red urchins. 

## Help to contextualize any differences you observe - for example, what is the actual and % difference in mean urchin diameter between the species?
- actual difference is 2.31, however percent difference is 0.011% (that is, the mean/sample size of red urchin minus purple urchin * 100)
```{r, message = FALSE, warning=FALSE}
red_urchin_diff = 6.776489/27408
purple_urchin_diff = 4.469797/31805
red_purple_diff = (red_urchin_diff - purple_urchin_diff) *100
red_purple_diff
```

## Explore purple urchin counts by site and year
```{r, message = FALSE, warning=FALSE}
urchins_clean <- urchins_clean %>% 
  mutate(date = mdy(date))

class(urchins_clean$date)

purple_urchins <- urchins_clean %>% 
  filter(common_name %in% c("Purple Urchin")) %>%
  mutate(Date = ymd(date)) %>% 
  mutate(Year = year(Date)) %>% 
  mutate(site_full = case_when(
    site == c("NAPL") ~ "Naples Reef", 
    site == c("AQUE") ~ "Arroyo Quemado Reef",
    site == c("CARP") ~ "Carpinteria Reef",
    site == c("MOHK") ~ "Mohawk Reef",
    site == c("IVEE") ~ "Isla Vista Reef"))

urchins_sum <- purple_urchins %>% 
  group_by(site_full, Year) %>% 
  count(site_full)
```


## Make an exploratory graph of counts of purple urchins recorded at each site over time for the years included in the study
```{r, message = FALSE, warning=FALSE}
ggplot(data=urchins_sum, aes(x = Year, y = n)) + 
  geom_line(aes(col = site_full)) +
  geom_point(aes(col = site_full))

ggplot(data = urchins_sum, aes(x = Year, y = n))+
  geom_col(aes(color = site_full)) +
  facet_wrap(~site_full)
```

## END